On the performance of the MAC model

### 2018-10-30 2018_nips_vigil/mac.tex
Wrapped figure in text. Modified equations to make it more readable.

The MAC network~\cite{hudsonManning18} is a recurrent model that performs reasoning 
using multiple steps, where each step involves sequentially analyzing a part of the question following which attention on the image is changed suitably.
At the end, a suitable answer is given based on this composite reasoning.
To improve the accuracy of prediction, both the questions and images
are preprocessed suitably. For the question,  a word-embedding is first used to
transformed the question into a sequence of vectors which is then passed through
a bidirectional LSTM to produce a sequence of hidden state pairs. Each such pair 
corresponds to a unique word in the question and is called the 
\emph{contextual embedding} for that word.
Similarly each image is initially processed using a pre-trained network (using ResNet) followed by two trainable CNN layers.\\
The core of the network is a sequence of recurrent MAC cells. A MAC cell (Figure 1) is performing one reasoning step. At every step the network is driving the attention on one part of the question and the corresponding part of the image. This operation is made possible by the combination of 3 units working together, a control unit, a read and a write units. The control unit is updating the control state and drives the attention over the question. Given this control state, the read unit then retrieves information from the memory and drives the attention over the image.
Finally, the write unit retrieves this visual information and updates the memory state.
The last part of the network is the output unit. The output unit is simply a classifier that process the concatenation of the question representation and the final memory to produce a final answer.\\
By going through this recurrent process, the network performs a reasoning strategy that learns concepts and relations between them. MAC is performing state of the art accuracy on CLEVR and CLEVR Humans datasets.
In contrast to modular approaches (Andreas et al., 2016a;b; Hu et al., 2017; Johnson et al., 2017b), MAC is fully
differentiable and does not require additional supervision. It is making use of a single computational cell chained in sequence rather than a collection of custom modules deployed in a rigid tree structure.
In contrast to augmented CNN approaches (Santoro et al., 2017; Perez et al., 2017),  the MAC model provides an ability for relational reasoning with better generalization capacity, higher
computational efficiency and enhanced transparency.

\begin{figure}[htbp]
	\centering
	\includegraphics[width=0.7\textwidth]{img/mac_cell.pdf}
	\caption{The Mac Cell. The MAC recurrent cell consists of a control unit, read unit, and write unit, that operate over dual control and memory hidden states. }
	\label{fig:mac_cell}
\end{figure}

>>>>

The MAC network~\cite{hudsonManning18} is a recurrent model that performs reasoning 
using multiple steps, where each step involves sequentially analyzing a part of the question following which attention on the image is changed suitably.
At the end, a suitable answer is given based on this composite reasoning.
To improve the accuracy of prediction, both the questions and images
are preprocessed suitably. For the question,  a word-embedding is first used to
transformed the question into a sequence of vectors which is then passed through
a bidirectional LSTM to produce a sequence of hidden state pairs. Each such pair 
corresponds to a unique word in the question and is called the 
\emph{contextual embedding} for that word.
Similarly each image is initially processed using a pre-trained network (using ResNet) followed by two trainable CNN layers.

\begin{wrapfigure}{r}[5pt]{0.4\textwidth}
	\vspace{-15pt}
	\centering
	\includegraphics[width=\textwidth]{img/mac_cell.pdf}
	\caption{The MAC cell~\cite{hudsonManning18}}
	\label{fig:mac_cell}
	\vspace{-5pt}
\end{wrapfigure}

The core of the network is a sequence of recurrent MAC cells. A MAC cell (Figure 1) is performing one reasoning step. At every step the network is driving the attention on one part of the question and the corresponding part of the image. This operation is made possible by the combination of 3 units working together, a control unit, a read and a write units. The control unit is updating the control state and drives the attention over the question. Given this control state, the read unit then retrieves information from the memory and drives the attention over the image.
Finally, the write unit retrieves this visual information and updates the memory state.
The last part of the network is the output unit. The output unit is simply a classifier that process the concatenation of the question representation and the final memory to produce a final answer.

By going through this recurrent process, the network performs a reasoning strategy that learns concepts and relations between them. MAC is performing state of the art accuracy on CLEVR and CLEVR Humans datasets.
In contrast to modular approaches (Andreas et al., 2016a;b; Hu et al., 2017; Johnson et al., 2017b), MAC is fully
differentiable and does not require additional supervision. It is making use of a single computational cell chained in sequence rather than a collection of custom modules deployed in a rigid tree structure.
In contrast to augmented CNN approaches (Santoro et al., 2017; Perez et al., 2017),  the MAC model provides an ability for relational reasoning with better generalization capacity, higher
computational efficiency and enhanced transparency.

%\begin{figure}[htbp]
%	\centering
%	\includegraphics[width=0.4\textwidth]{img/mac_cell.pdf}
%	\caption{The MAC cell~\cite{hudsonManning18}}
%	\label{fig:mac_cell}
%\end{figure}

### 2018-10-30 2018_nips_vigil/smac.tex
Wrapped figure in text. Modified equations to make it more readable.

\subsection*{Control Unit}

\begin{align*}
	&q_i = U_i^{[d \times 2d]} q + b_i^{[d]}  \tag{c0} \\
	\cline{1-2}
	&cq_i = W_{\textsf{cq}}^{[d \times d]} c_{i-1} + q_i  \tag{c1} \\
	&ca_{is} = W_{\textsf{ca}}^{[1 \times d]} (cq_i \odot \mathbf{cw}_s)  \tag{c2.1}\\
	&cv_{is} = \textrm{softmax}(ca_{is}) \tag{c2.2}\\
	&\mathbf{c}_i = \sum_s cv_{is} \, \mathbf{cw}_s  \tag{c2.3}
\end{align*}
Only eqn.~(c0) above is \emph{position dependent}

\subsection*{Read Unit}

\begin{align*}
	&\mathbf{\tk}_{hw} = U_{\textsf{proj}}^{[d \times d]} \mathbf{k}_{hw} + b_{\textsf{proj}}^{[d]} \tag{r0} \\
	\cline{1-2}
	&I_{ihw} = \mathbf{m}_{i-1} \odot \mathbf{\tk}_{hw} \tag{r1} \\
	&I'_{ihw} = \mathbf{\tk}_{hw}  + W_{\textsf{I}}^{[d \times d]} I_{ihw} + b_{\textsf{I}}^{[d]}  \tag{r2} \\
	&ra_{ihw} = W_{\textsf{ra}}^{[1 \times d]} (\mathbf{c}_i \odot I'_{ihw})  \tag{r3.1}\\
	&rv_{ihw} = \textrm{softmax}(ra_{ihw}) \tag{r3.2}\\
	&\mathbf{r}_i = \sum_s rv_{ihw} \, \mathbf{\tk}_{hw}  \tag{r3.3}
\end{align*}

Note that $\mathbf{\tk}_{hw}$ can be \emph{precomputed} before the reasoning steps.

\subsection*{Write Unit}

\begin{equation}
\mathbf{m}_i = W_{\textsf{m}}^{[d \times d]} \mathbf{r}_i + b_{\textsf{m}}^{[d]} \tag{w1}
\end{equation}

>>>>

In the description below, the original MAC cell equations are shown on the \emph{left}
while our simplified equations are shown (in color) on the {\color{Plum} \emph{right}}.
(The equation numbering is the same as in~\cite{hudsonManning18}.)

\noindent\textit{Notation:} The index/position $i$ below denotes the reasoning step. 
Unless mentioned otherwise, the weights/biases are invariant w.r.t $i$, 
i.e., shared across all the reasoning steps.

\noindent\textbf{Control unit:} 
For both models, in the control unit, the question $q$ is first transformed in each step of 
the reasoning using a \emph{position-aware}~\cite{hudsonManning18} 
linear layer: $q_i = U_i^{[d \times 2d]} q + b_i^{[d]}$.

\begin{multicols}{2}
	\noindent
	\begin{align*}
	&cq_i = W_{cq}^{[d \times 2d]} [c_{i-1}, q_i] + b_{cq}^{[d]}  \tag{c1} \\
	&ca_{is} = W_{ca}^{[1 \times d]} (cq_i \odot \cw_s) + b_{ca}^{[1]}
	\tag{c2.1}\\
	&cv_{is} = \textrm{softmax}(ca_{is}) \tag{c2.2}\\
	&\cc_i = \sum_s cv_{is} \, \cw_s  \tag{c2.3}
	\end{align*}
	\columnbreak
	{\color{Plum}
	\begin{align*}
	&cq_i = W_{cq}^{[d \times d]} c_{i-1} + q_i  \tag{c1} \\
	&ca_{is} = W_{ca}^{[1 \times d]} (cq_i \odot \cw_s)  \tag{c2.1}\\
	&cv_{is} = \textrm{softmax}(ca_{is}) \tag{c2.2}\\
	&\cc_i = \sum_s cv_{is} \, \cw_s  \tag{c2.3}
    \end{align*}}
\end{multicols}

\noindent\textbf{Read and write units:}
For both models below, $\kb_{hw}$ denotes the \emph{knowledge base}~\cite{hudsonManning18}
consisting of 3D tensors with dimension $H \times W \times d$ for each image
in the data set. The difference in computing $\kb_{hw}$ is that the simplified MAC uses
an additional linear layer on top of the 2 CNN layers present in the original MAC network.
This circumvents the need for performing additional linear transformations on the 
knowledge base, as is done by the original MAC network in \emph{each} reasoning step.

\begin{multicols}{2}
	\noindent
	\begin{align*}
	&I_{ihw} = (W_{m}^{[d \times d]} m_{i-1} + b_{m}^{[d]}) \\
	           & \qquad \quad \odot (W_{k}^{[d \times d]} \kb_{hw} + b_{m}^{[d]}) \tag{r1} \\
	&I'_{ihw} =  W_{I'}^{[d \times 2d]} [I_{ihw},\kb_{hw}]  + b_{I'}^{[d]}  \tag{r2} \\
	&ra_{ihw} = W_{ra}^{[1 \times d]} (\cc_i \odot I'_{ihw}) + b_{ra}^{[1]} \tag{r3.1}\\
	&rv_{ihw} = \textrm{softmax}(ra_{ihw}) \tag{r3.2}\\
	&\rr_i = \sum_s rv_{ihw} \, \kb_{hw}  \tag{r3.3}\\
	&\mem_i = W_{rm}^{[d \times d]} [\rr_i, \mem_{i-1}]  + b_{rm}^{[d]} \tag{w1}	
	\end{align*}
	\columnbreak
	{\color{Plum}
	\begin{align*}
	&I_{ihw} = \mem_{i-1} \odot \kb_{hw} \tag{r1} \\ \\
	&I'_{ihw} = W_{I'}^{[d \times d]} I_{ihw} + b_{I'}^{[d]} + \kb_{hw} \tag{r2} \\
	&ra_{ihw} = W_{ra}^{[1 \times d]} (\cc_i \odot I'_{ihw})  \tag{r3.1}\\
	&rv_{ihw} = \textrm{softmax}(ra_{ihw}) \tag{r3.2}\\
	&\rr_i = \sum_s rv_{ihw} \, \kb_{hw}  \tag{r3.3}\\
	&\mem_i = W_{rm}^{[d \times d]} \rr_i + b_{rm}^{[d]} \tag{w1}
	\end{align*}}
\end{multicols}
